Under-ice acoustic navigation using real-time model-aided range estimation

### 2022-02-21 rebuttal/rebuttal-letter/trunk-rebuttal.tex

\input{../p-rebuttal-letter.tex}

% table of contents
\renewcommand{\contentsname}{}
\begingroup
\let\pagebreak\relax
\setcounter{tocdepth}{1}
\tableofcontents
\endgroup

\section*{Cover Letter}
Dear Editors,

We thank the reviewers and editor for their thoughtful feedback and comments in improving this work.
We have refashioned the abstract and introduction to better elucidate the motivation for this work as a model-aided component for LBL navigation in total under-ice conditions, which could be further generalized for any acoustic environment.
Based on several suggestions, we have restructured the manuscript and separated the real-time from the post-processed range estimation results.
Our original revision aimed to contextualize the range estimation results as a proxy for navigation.
Given several points from the reviewers and editor about the lack of this transferability, we have added AUV post-processed navigation results as well.
The real-time results are also covered but have no GNSS ground truth to compare to.

@@

\begin{enumerate}
	\item Positioning is the ability to accurately and precisely determine one’s location referenced to a standard geodetic system;
	\item Navigation is the ability to determine current and desired position (relative or absolute) and apply corrections to course, orientation, and speed to attain a desired position anywhere in the domain of concern; 
	\item Timing is the ability to acquire and maintain accurate and precise time anywhere in the domain of interest within user-defined timeliness parameters; it also includes time transfer.
\end{enumerate}

Navigation is real time and necessarily depends on positioning, and positioning depends on timing.
\end{quote}

Given the tension of evaluating our travel time to range conversion on events between LBL beacons in real time and post-processing, we further delineate that, in this work, positioning is in real time and re-positioning is post-processed.
We have also added re-positioning trilateration results, processed with only the available field data and algorithms that could be performed in real time, to emulate navigation performance.

The rebuttal letter is organized with sections delineated by the review document and email. We look forward to your comments and the opportunity to continue improving this manuscript for potential publication in JASA. Thank you again for your time and consideration.

@@

	\ebline{This paper demonstrates how a positioning system over \texttt{\small $\sim$}1.5-3 km of range achieves 11 m of positioning uncertainty. The work demonstrates how a single group velocity calculated from acoustic rays predicted by BELLHOP can be used to estimate ranges to be used in a LBL network for positioning. The data presented are from transceivers on fixed ice-tethered buoys but the system could also be implemented on an autonomous underwater vehicle.}

	This summary captures key results presented in the paper. However, we note that the system was successfully deployed on an underwater vehicle during ICEX20. The results from this work are specifically limited to the methodology behind calculating a single group velocity, which we attempt to quantify through beacon-to-beacon communications with known GNSS timestamps.

	\subsection*{Overall Comments}

	\reviewer{}{The claims made in the abstract need to be better supported by the data presented or else tempered a bit. What is demonstrated here is not real-time navigation on a vehicle, but calculation of a handful of short fixed ranges from ice-tethered buoys. The limited data presented here begs more discussion with regards to how applicable the algorithm would be in normal vehicle operation and to navigation of the vehicle, as there were limited ranges between the beacons. What depths does the vehicle typically operate at and what is the maximum range from the beacons?  The abstract stresses real-time results, but the real-time results are just 3 paragraphs, and the paper mostly concentrates on post-processing results.} \label{1dot1}

	We have adjusted the claims made in the abstract (and throughout the paper) and hope we have presented data to better support them as well. 
	While we initially aimed to contextualize the calculation of a handful of short fixed ranges between ice-tethered buoys as the proxy for navigation, we agree with the reviewer that this is a severe shortcoming. Accordingly, we added the same analysis for the AUV run in ICEX20.

	We have also addressed valid questions about transferability to other vehicle operations in the discussion (line \ref{v1:1.1}).

	\reviewer{}{The hypothesis is stated in the discussion section lines 515 and following: “We hypothesize and validate that the embedded stochastic prediction of a single group velocity is a smoothly varying function of range, source, and receiver depth pairings as well as multipath structure.”  I am not convinced that the few short fixed ranges presented here validate this hypothesis.  Lines 248 and following indicate how the system should track a continuously changing group velocity, but this is not demonstrated with the few fixed ranges presented here.  Also, it would be good to state this hypothesis more clearly earlier in the manuscript.}

	We agree that a few short fixed ranges do not validate this hypothesis. With the overarching motivation and design of LBL navigation, however, we have recast this hypothesis as an assumption (line \ref{v1:2.10}).

@@

	The error metrics presented here should be specifically compared with other positioning systems or algorithms over similar spatial scales that do not integrate the ray propagation physics.

	Various ranging and positioning methods are referenced, but the narrative jumps between different types of AUVs and implementations and between scales of a few kilometers and hundreds of kilometers. This should be discussed in terms of range, vehicle, and propagation conditions.

	Line 176 mentions that the SSP and the physics of ray refraction are often overlooked in positioning algorithms; this is, I presume, referring to real-time results or short-range results, but doesn’t specifically say this.}

	We believe we have tightened up the introduction and conclusion sections by focusing on this work as a model-aided augmentation to an LBL system.
	Then we specifically talk about AUV-LBL systems deployed in under-ice conditions.
	This focus presents the narrative more linearly, instead of it feeling like it jumps between different types of deployments.

	The error metrics are presented as RMS m. Other positioning systems at similar scales are now covered in the introduction (line \ref{v1:1.3a}). We have also added an isovelocity case for fairer comparison to other smaller spatial scale AUV deployments and the proposed method in this manuscript.

	Thank you for the suggestion to discuss in terms of range, vehicle, and propagation conditions. We have not done this explicitly but taken care to illustrate the differences, and what parts we are comparing.

	Lastly, yes, we did mean real-time navigation, but have removed the line in question.

@@

	\reviewer{}{Related to the point above, the introduction should make clear what type of vehicles the proposed method of navigation is applicable to and over what ranges. The authors do start to do this when they note that a shallow receiver depth enables the estimation method (line 274), but should more directly specify under what conditions the proposed algorithm is applicable.  In the Discussion section, there is an implication that this method of using a single group velocity could be applied to mesoscale operations but should acknowledge that propagation at these larger scales results in arrivals that are spread over several seconds including multiple modal arrivals with varying group velocities.  This method would presumably result in errors much higher than 11m….how might that scale?}

	{\color{blue} COME BACK TO THIS.}

	\reviewer{}{I understand what the authors are referencing when they use the term “group velocity” but I think this term should be used a bit more carefully.  Group velocities are often associated with modal propagation in which each mode has a distinct group velocity.  The goal in this work is to determine a single group velocity by taking a power weighted average of several different ray arrivals to derive range.  This is addressed somewhat in lines 395 and following, but I think it would be appropriate to expand on this discussion and move it up.  Lines 235 and following state “the variation of horizontal group velocity from any source-receiver pair is the fundamental challenge to implement GPS-like navigation in an LBL navigation paradigm…” this statement could use some clarification.  Also, perhaps change the axis on Figure 5 to read something like ‘implied group velocity.”} \label{1dot5}

	Thank you, these are all excellent suggestions. Instead of ``group velocity'' we use a combination of ``effective sound speed'', noted by $c$. The statement in question has been rewritten to talk about a single sound speed compensating for ray refraction and reflection, and is on line \ref{v1:1.5}.

	\reviewer{}{As written, this manuscript seems more suited to an ocean engineering journal.  For a JASA audience, this should be more focused on the acoustics.  Perhaps I missed it, but I didn’t see that the manuscript even mentioned the frequencies or waveforms that were used in the experiment.  I’d suggest, for example, less emphasis on which options were toggled on/off or triggered (e.g. line 186, 196, 241, 254 ) and more on, for example, why specific depths were chosen and how they provide better coverage due to the acoustic shadow zone (e.g., Line 155), or how the specific rays were eliminated by the filtering (Line 232). Lines 370-372 state “The increased error for these reciprocal transmission paths is most likely driven by the computational artifacts encountered when propagating through the steep sound speed gradients of the lens and through the shadow zone.“ The authors might consider providing more information on these artifacts.}\label{1dot6}

	Agreed on all points. The frequencies are now included in the ICNN description (line \ref{v1:1.6a}). The discussion of which options were toggled on and off has been removed in favor of a in depth discussion of shadow zone, computational artifacts, and eigenray filtering in Section III.C, starting line \ref{v1:1.6b}.

@@

	\reviewer{}{There seems to be a somewhat outsized discussion of GPS drift, but I didn’t see how the GPS uncertainty was directly applied to the uncertainty values presented in the paper.  I’d suggest condensing this section.}

	This section has been significantly condensed and is now Section V.C, starting line \ref{v1:1.7}. The GPS uncertainty is now directly compared to the uncertainty from beacon-to-beacon re-positioning and AUV re-navigation.

@@

	\subsection*{Specific Comments}

	\lreviewer{line 226}{v1:1.8}
	{In the section describing post-processing analysis, the authors bring up that the processing was designed to mirror information available on a submarine – modeled data, historical data, and in situ data.  This is a nice rationale for the analysis, and I’d suggest highlighting this in the introduction.}

	Thank you. This is now included in the ICEX20 experiment description, Section II.

@@

	\lreviewer{}{}{How long was the experiment from start to finish, and how much did the sound speed profile change over the duration of the experiment?}

	This is a great question! The experiment was four days. The last few CTDs showed a significantly weaker duct than those at the beginning. The local sound speed maxima went from 1447 m/s at 40 m in depth to 1442 m/s at 60 m in depth. Related work on the environmental adaptivity has just been accepted in IEEE JOE.

@@

	\lreviewer{}{}{What is the magnitude of the uncertainty in the position of the transceivers relative to the GPS positions measured at the surface? Presumably the transceivers particularly at 90 m depth could move from the lat/lon position of the GPS receivers at the surface.}

	We assume negligible uncertainty in the position of the modems relative to the surface expression given that the ice drift rate was unusually low, roughly 0.7 cm/s. This is now included on line \ref{v1:1.10}.

@@

	\lreviewer{lines 40-42}{v1:1.11}{Note ranging via one way travel times that have been used prior to the advent of the micro-modem, for example RAFOS.}

	Thank you. We have included this point and make the distinction that RAFOS floats are re-positioned.

@@

	\lreviewer{line 192}{v1:1.12}{Clarify what you mean when you state that the transmit and receive layers are independent? Also, the discussions of physical and virtual layers are confusing. I think everything is there, but reorganization would make this more clear. Consider moving up the paragraph starting at line 203.}

	We have removed the jargon of the physical and virtual layers in favor of a more streamlined explanation. The transmit and receive layers are activated independently.

	\lreviewer{line 205}{v1:1.13}{What do you mean that the coordinate system is range independent?}
	We meant that the analysis of beacon-to-beacon connections is presented in r/z, not x/y/z. We found this framing awkward as it is more common to say the propagation has a depth dependent sound speed profile with range dependent bathymetry.

@@

	\lreviewer{line 222}{v1:1.14}{Could you provide more information on the baseline and weighted SSPs. Where did the baseline profile come from?}
	\label{1dot14}

	In anticipation of ICEX20, all of the ITP data for the Beaufort Sea was assimilated. For operation reasons, we designed a system to transmit sound speed updates through digital acoustic message. This system relied on a simple basis function representation of the sound speed and its perturbations. Given the wide range of observed ducted conditions across all ITP data, several basis sets were created, corresponding to March of different years. During ICEX20, the data from Mar 2013 was found to be the closest match for observed conditions. The baseline is the mean sound speed from Mar 2013. The weighted is the basis set projection for the CTD taken during ICEX20.

@@

	\lreviewer{line 256}{v1:1.15}{How is the group velocity uncertainty calculated?}

	We meant standard deviation, and have changed this accordingly. It is equation 2.

@@

	\lreviewer{line 231}{v1:1.16}{Quantify the travel-time proximity to recorded data that was used to filter the eigenrays.}

	It was a one millisecond filter to cluster distinct arrivals.

@@

	\lreviewer{lines 260, 303-305}{v1:1.17}{Please provide more information on the BELLHOP implementation.   Line 260 indicates a sparse grid.  From lines 303-305 do I understand that the grid is 1m in range and 2 m in depth?}

	This was not clear. We meant sparse as in not full-field, which we agree is misleading. The grid is 121 points with 1 m spacing in range and 2 spacing m in depth.

@@

	\lreviewer{line 309}{v1:1.18}{Suggest re-writing this sentence for clarity: “As expected, the algorithm is generally overestimating range as it resolves a direct path that does not represent the actual arrival.”}

	This has been rewritten as ``The algorithm generally overestimates pseudoranges because it resolves the effective sound speed for the most direct path.''

@@

	\lreviewer{line 364}{v1:1.19}{Could use a figure reference.}

	We have added a figure reference.

	\lreviewer{line 562}{v1:1.20}{I don’t agree that many approaches to underwater navigation combine it with acoustic tomography.}

	We agree, ``many'' misdirects the current state of the art. The benefit of concurrent navigation and tomography has been discussed, but it remains to be used as a standard technique.

	\lreviewer{Figure 1}{v1:fig:vehicleRecovery}{Figure 1 is a nice picture showing the vehicle, which motivates the work presented here, but was not used for collection of any of the data presented.  As it’s not pertinent to the data presented here, I’d suggest removing this figure or at least combining this figure with Figure 2.}

	Thank you. We have kept this image but added the vehicle track, as it is necessary for the re-navigation analysis as mentioned in Comment \ref{1dot1}.

	\lreviewer{Figure 3}{v1:fig:overview}{Figure 3 tallies the number of successful events, and I think the use of the directional arrows communicates the data well.  How many events are successful in comparison with those that are considered unsuccessful and what generally causes an event to be deemed unsuccessful?}

	Thank you. We have added a successful decoding percentage, defined as the ratio of decoded over detected messages. We have also added the total number of transmissions, so one can see the ratio of detected over transmitted. An event is deemed unsuccessful given any error in the data frame, even if the error has nothing to do with the information necessary for positioning.

	\lreviewer{Figure 5}{v1:fig:gvel-post}{What is the point labeled u=751 m/s that appears to fall around 1424 m/s?}

	That point is an effective sound speed prediction dominated by a bottom bounce. The figure has been changed to include all source depths. Keeping this point in would visually compress the other data. However, it is now discussed narratively, in line \ref{v1:1.23}.

@@

	\lreviewer{Table II}{}{Might this information be portrayed better by a box plot?}

	This table described pseudorange statistics. Given that these are no longer serving as a proxy for navigation, it has been removed. The statistics for re-positioning and re-navigation are now included as histogram-box plots (Figs 11 and 12).

@@

	Thank you. We have standardized this to Topside.

	\lreviewer{lines 311, 359}{}{'Data' should be plural.}
	Thank you. This has been corrected.

	\lreviewer{Figure 6}{v1:fig:rangeError}{Suggest including receiver depth along with source depth in the title as they are both important in distinguishing between the cases.}

	The source depths have been moved from the column titles to the individual sub-panels.

	\lreviewer{Table I}{}{Table I indicates direction errors are positive. Please note the sign convention in the caption.}

	We have removed the table but discuss this in the text at line \ref{v1:1.28}, and have corrected the convention.

	\lreviewer{Table II}{}{‘range estimation error metrics’ rather than ‘range estimation metrics’?}
	Agreed. While we have removed the table, we have adopted this language.

	\lreviewer{Table II and III}{}{Suggest removing ‘algorithm’ label.  Also substitute ‘MBC’ and ‘NBC’ for minimal and nearest to match the text.}
	Agreed. While we have removed the table in favor of histogram/box plot, we have adopted this language.

	\lreviewer{}{}{There are several figures and tables which are not referenced properly (Line 305, 391, 411)}
	Thank you. This has been corrected and we apologize for the typographical mistakes.

@@

%% ------- COMMENTS FROM REVIEWER TWO ------ %%

\section{Reviewer Two}

	\lreviewer{line 7}{v1:2.1}{'Beaufort Lens' This is not yet a scientific term, suggest finding descriptive terms to use. What comprises the 'lens'? The Pacific Summer Water is the warm layer.}\label{2dot1}

	We have removed the ``Beaufort Lens'' language from the abstract as it may not be self explanatory. However, the Beaufort Lens has been used in the literature (Litvak 2015, Schmidt and Schneider 2016) and in JASA/POMA specifically (Poulsen et al 2016, Chen et al 2019).

	We think of the lens as any feature that introduces significant refraction, such that the term Beaufort Lens is a convenient shorthand for the spatio-temporal variability of the warm water intrusion provided by the Pacific Summer Water Layer that creates two ducts. We prefer this distinction as the lens exists in the oceanographic space but the ducts are acoustic features dependent on source depth and frequency. With this jargon, the lens is susceptible to aberrations from eddies or filaments in the same way that the duct is susceptible to leaking.

	For clarity, we explictly define the Beaufort Lens in the manuscript as discussed here at line \ref{v1:2.1a}.

@@

	\lreviewer{line 10}{v1:2.2}{'GPS aided', do you mean GPS-navigated?}
	We have changed this to ``GPS-linked'' because navigation implies self-movement.

	\lreviewer{line 12}{v1:2.3}{'rival GPS' is a fairly bold statement when GPS is used as ground truth.} \label{2dot3}

	We meant ``rival GPS'' in that the downstream accuracy of the acoustic range estimation method is within the precision of GPS. However, we have amended this framing to say that our best performing acoustic method does not introduce significant additional error than is already documented by GPS puck precision, effectively extending single GPS performance into the shallow water column for a small volume of interest. Given more beacons, there is potential to reduce uncertainty in the same way that differential GPS works. While this refrain is used multiple times in the revised manuscript, please see Section V in particular.

@@

	\lreviewer{line 224}{v1:2.7}{why did a human have to chose an optimal set of weights? Why not just use the measured SSP, or an average of a few?}

	We have removed this statement, as it is not relevant to the findings of this work. However, the navigation solution existed in tandem with an interactive tactical decision aid (TDA) operated by a Navy submarine officer. A lightweight compression of the sound speed profile was needed to facilitate its update via digital acoustic message. Work about this TDA has been recently accepted for publication in IEEE JOE.

	\lreviewer{line 231}{v1:2.8}{'abundance'. Is there a more precise way to say this? Bellhop does what it is programmed to…}

	Yes, thank you. BELLHOP finds all possible eigenrays given the prescribed launch angles whereas only a few or less (sometimes none) are valid for the travel time of interest.

@@

	\lreviewer{line 241}{v1:2.9}{'modem triggering time', do you mean 'detected arrival time'?}
	Yes. We have changed this accordingly.

### 2022-02-21 manuscript/main.tex
added tags + more complete answers to rebuttal letter

% More about LBL navigation
While RAFOS floats championed one way ranging for re-positioning \citep{Rossby1986,duda_evaluation_2006}, the ability to do so for navigation was facilitated by \llabel{1.11} the advent of the WHOI Micro-Modem \citep{Singh2006} and synchronized clocks \citep{rypkema_one-way_2017}.
AUV navigation efforts have achieved root mean square (RMS) localization error on the order of tens of meters relative to GNSS surface position over less than ten kilometers in shallow \citep{Eustice2007,Claus2018,kepper_mems_2017} and deep water \citep{Kunz2008,jakuba_long-baseline_2008,Webster2009}.
However, these efforts all used a nominal sound speed for travel time conversion and the vehicles were limited to shallower isovelocity regimes.

@@

Given that a lens introduces significant ray refraction, the Beaufort Lens is a shorthand for the spatio-temporal variability of the local temperature and sound speed maxima generally around 50 to 60 m in depth.
A neutrally buoyant layer of warm Pacific Summer Water creates a unique double ducted environment \textemdash the upper duct degrades signal coherence due to intensified ice interaction and the lower duct effectively traps sound for long range propagation \citep{Poulsen2016}. 
Modeling output \citep{duda_long-range_2019,duda_effects_2021} and experimental observations \citep{Ballard2020,Bhatt2021} suggest that, in the Beaufort Sea, the duct is persistent and widespread but not necessarily continuous; it and its acoustic effects can be non-existent, minimal, or drastic.
Transmissions in the upper duct, between the surface ice and the lens, experience minimal attenuation but degrade in signal coherence with repeated reflections under the ice.
In lower duct, between the lens and its conjugate depth in the Atlantic water, roughly 200 m, sound above 350 Hz is trapped near losslessly for long range propagation \citep{Poulsen2016}.

@@

The ICNN is comprised of four ice buoys in a loose rectangle, roughly 2 km away from a central ice camp with a topside computer, as shown in Fig.~\ref{fig:icnnOverview}.
Each ice buoy is outfitted with a Garmin GPS 18x, with a pulse-per-second rising edge aligned to 1 microsecond and a spec sheet accuracy of 3 m, 95\% of the time.
The AUV and each ice buoy are outfitted with a WHOI Micro-Modem \citep{Singh2006}, with a four-element receiver array, a single transmitter, and one-tenth of a millisecond resolution.
Acoustic messages were sent with a 10 kHz carrier frequency, 5 kHz bandwidth, and phase-shift keying (PSK)modulation on a time-division multiple access schedule with a thirty-second cycle, giving room for two-way communication throughout the mission volume.
The receive and transmit elements were split between shallow and deeper depths\textemdash 30 and 90 m\textemdash to provide better coverage across the shadow zone.
While each buoy only has one transmit depth, all buoys have both receive depths but the active receive layer is consistent across all buoys. \llabel{1.12}
The design of the ICNN enables a self-adapting network to transmit and receive at the optimal depth to maintain contact with the AUV \citep{schneider_self-adapting_2020}.
The buoys do not encompass the full horizontal range of the vehicle but are positioned to minimize overlap in trilateration for spherical positioning \citep{deffenbaugh1996posit}.

@@

Because the vehicle's navigation solution during a mission can only be evaluated on the basis of the error estimates sent, a sister experiment for validating the real-time ranging approach was implemented.
Ice buoy modems were run as ``virtual vehicles'' at a fixed depth, receiving position updates from the other beacons as well as a camp site modem lowered to 20 m. \llabel{1.13}
Fig. \ref{fig:overview} shows successful events sorted by source depth.
In this analysis, we assume there is insignificant displacement between the GNSS puck surface expression and subsurface modem; this is supported by unusually low observed ice drift rates, just 0.7 cm/s on average throughout the mission.

@@

The fundamental challenge to implement GNSS-like navigation, especially in an acoustically complex propagation environment, is characterizing a single sound speed to compensate for the effects of ray refraction and reflection.
The use of the acoustic modem network for tracking relies on the accurate estimates of travel times between the submerged platform and LBL beacons, supported by clock synchronization and a pre-determined scheduling of acoustic events.
\llabel{2.9} For the Beaufort Lens in particular, the strong multipath effects make it virtually impossible to deterministically predict the modem's detected arrival time.

@@

Fig. \ref{fig:range-error-insitu} shows the range error boundary for both SSP inputs used in ICEX20.
A promising sign that the MBC method adapts sound speed somewhat intelligently is the lack of error growth as travel time increases.
The baseline SSP (n=243 events) has an absolute pseudorange error of 11.38 $\pm$ 4.23 m; the weighted SSP (n=568), 11.36 $\pm$ 8.12 m.
The discrepancy between these two is largely due to outlier events only contained in the weighted SSP set.
Where there is overlap between sound speed conditions used for the real-time MBC, the pseudorange error difference is no more than a few meters.
The overarching results show that sound speed estimates derived from eigenrays for a local grid, as opposed to a singular point, are accurate enough to support vehicle navigation.
While the NBC looks for just the least complex multipath, the high density of launch angles almost always guarantees a direct path.
Nonetheless, the consistent overestimation of pseudorange invites further analysis into acoustic arrival matching.

\subsection{Eigenray identification for beacon-to-beacon events} \label{sec:eigenrays}

@@

The goal of the MBC algorithm was to provide a reliable, physically intuitive interpretation of the acoustic propagation without taking on the additional burden of regularly identifying specific paths that may connect any given source-receiver pair in the network.
\llabel{2.14}
While it was unlikely to resolve multipath arrivals that triggered successful modem detection, an isovelocity approach would have provided no adaptivity against source and receiver depth differences.
Its performance was adequate for vehicle navigation and would have likely sufficed if it were not for the prominence of the duct observed relative that of other model and data products.

@@

The various sound speed inputs\textemdash isovelocity aside\textemdash not only modify the predicted effective sound speed, as seen by the colored vertical offsets, but often classify a distinct number of bounces.
HYCOM sees the most direct and one bounce multipath structures, lending a bias for faster speeds; the weighted SSP sees the most double and triple bounces, favoring slower speeds; the baseline sound speed exists in between.
Very rarely is the multipath structure classified as a direct path, where the MBC and NBC would prediction overlap.
In fact, the higher the multipath classification, the more accurate the sound speed prediction is, likely driven by a tighter or even sparser bundle of rays.
\llabel{1.19} Discontinuities in multipath classification provide initial evidence for its importance to a smoothly varying group velocity, as shown in the cluster of 30 to 30 m transmissions, where HYCOM jumps from one to two classified bounces amidst the baseline SSP and weighted SSP smoothly increasing while consistently seeing two and three classified bounces, respectively.
Of course, the prediction deteriorates with cross-layer transmissions across the duct, but not to the same degree at which eigenrays could not be found for the weighted SSP in section \ref{sec:eigenrays}.
The evidence suggests that the grid based method provides a useful amount of redundancy to resolve similar enough eigenrays.

It is useful to think about in what case the isovelocity\textemdash or any isovelocity framing\textemdash would have been appropriate.
The transmissions from shallow to shallow receiver would may have matched the default configuration of 1430 m/s.
The isovelocity contrived for this paper, 1441.8 m/s, best matches the transmissions from 90 to 90 m.
The one from \citet{Graupe2019}, 1450 m/s, would have had a systemic overestimation.
Given that implied sound speeds just for beacon-to-beacon events span 1420 to 1445 m/s, it is safe to say that a nominal sound speed would sacrifice pseudorange accuracy somewhere, and that an adaptive approach is necessary even for short range operations in the Beaufort Lens.

@@

There is one example that helpfully illustrates the improvement brought upon by bounce classification.
For transmissions between North and South at 30 m, the OWTT spread is 2.1958 to 2.1963 s; the GNSS-tracked distance is 3138.54 to 3140.87 m; and the implied effective sound speed is 1429.3 to 1430.1 m/s.
For these transmissions, the weighted SSP and the MBC approach produce a pseudorange error of -1491 m, as the effective sound speed is dominated by bottom bounce arrivals with much greater travel times.
The NBC approach categorizes this same record as a quadruple surface bounce, reducing the pseudorange error to less than a meter.
Comparatively, the NBC approach for HYCOM and the baseline SSP produce pseudorange errors of 8.30 and 2.39 m, respectively. 
There is strong evidence to suggest that the sound speed and multipath fidelity codependently improve localization accuracy.

@@

\subsection{Investigating potential GNSS noise}

>>>>

% More about LBL navigation
While RAFOS floats championed one way ranging for re-positioning \citep{Rossby1986,duda_evaluation_2006}, the ability to do so for navigation was facilitated by \llabel{1.11} the advent of the WHOI Micro-Modem \citep{Singh2006} and synchronized clocks \citep{rypkema_one-way_2017}.
\llabel{1.3a}
AUV navigation efforts have achieved root mean square (RMS) localization error on the order of tens of meters relative to GNSS surface position over less than ten kilometers in shallow \citep{Eustice2007,Claus2018,kepper_mems_2017} and deep water \citep{Kunz2008,jakuba_long-baseline_2008,Webster2009}.
However, these efforts all used a nominal sound speed for travel time conversion and the vehicles were limited to shallower isovelocity regimes.

@@

\llabel{2.1a}
Given that a lens introduces significant ray refraction, the Beaufort Lens is a shorthand for the spatio-temporal variability of the local temperature and sound speed maxima generally around 50 to 60 m in depth.
\llabel{2.4}
A neutrally buoyant layer of warm Pacific Summer Water creates a unique double ducted environment \textemdash the upper duct degrades signal coherence due to intensified ice interaction and the lower duct effectively traps sound for long range propagation \citep{Poulsen2016}. 
Modeling output \citep{duda_long-range_2019,duda_effects_2021} and experimental observations \citep{Ballard2020,Bhatt2021} suggest that, in the Beaufort Sea, the duct is persistent and widespread but not necessarily continuous; it and its acoustic effects can be non-existent, minimal, or drastic.
Transmissions in the upper duct, between the surface ice and the lens, experience minimal attenuation but degrade in signal coherence with repeated reflections under the ice.
In lower duct, between the lens and its conjugate depth in the Atlantic water, roughly 200 m, sound above 350 Hz is trapped near losslessly for long range propagation \citep{Poulsen2016}.

@@

The ICNN is comprised of four ice buoys in a loose rectangle, roughly 2 km away from a central ice camp with a topside computer, as shown in Fig.~\ref{fig:icnnOverview}.
Each ice buoy is outfitted with a Garmin GPS 18x, with a pulse-per-second rising edge aligned to 1 microsecond and a spec sheet accuracy of 3 m, 95\% of the time.
The AUV and each ice buoy are outfitted with a WHOI Micro-Modem \citep{Singh2006}, with a four-element receiver array, a single transmitter, and one-tenth of a millisecond resolution.
\llabel{1.6a} Acoustic messages were sent with a 10 kHz carrier frequency, 5 kHz bandwidth, and phase-shift keying (PSK) modulation on a time-division multiple access schedule with a thirty-second cycle, giving room for two-way communication throughout the mission volume.
The receive and transmit elements were split between shallow and deeper depths\textemdash 30 and 90 m\textemdash to provide better coverage across the shadow zone.
While each buoy only has one transmit depth, all buoys have both receive depths but the active receive layer is consistent across all buoys. \llabel{1.12}
The design of the ICNN enables a self-adapting network to transmit and receive at the optimal depth to maintain contact with the AUV \citep{schneider_self-adapting_2020}.
The buoys do not encompass the full horizontal range of the vehicle but are positioned to minimize overlap in trilateration for spherical positioning \citep{deffenbaugh1996posit}.

@@

Because the vehicle's navigation solution during a mission can only be evaluated on the basis of the error estimates sent, a sister experiment for validating the real-time ranging approach was implemented.
Ice buoy modems were run as ``virtual vehicles'' at a fixed depth, receiving position updates from the other beacons as well as a camp site modem lowered to 20 m. \llabel{1.13}
Fig. \ref{fig:overview} shows successful events sorted by source depth.
\llabel{1.10}
In this analysis, we assume there is insignificant displacement between the GNSS puck surface expression and subsurface modem; this is supported by unusually low observed ice drift rates, just 0.7 cm/s on average throughout the mission.

@@

\llabel{1.5} The fundamental challenge to implement GNSS-like navigation, especially in an acoustically complex propagation environment, is characterizing a single sound speed to compensate for the effects of ray refraction and reflection.
The use of the acoustic modem network for tracking relies on the accurate estimates of travel times between the submerged platform and LBL beacons, supported by clock synchronization and a pre-determined scheduling of acoustic events.
\llabel{2.9} For the Beaufort Lens in particular, the strong multipath effects make it virtually impossible to deterministically predict the modem's detected arrival time.

@@

Fig. \ref{fig:range-error-insitu} shows the range error boundary for both SSP inputs used in ICEX20.
A promising sign that the MBC method adapts sound speed somewhat intelligently is the lack of error growth as travel time increases.
\llabel{1.28}
The baseline SSP (n=243 events) has an absolute pseudorange error of 11.38 $\pm$ 4.23 m; the weighted SSP (n=568), 11.36 $\pm$ 8.12 m.
The discrepancy between these two is largely due to outlier events only contained in the weighted SSP set.
Where there is overlap between sound speed conditions used for the real-time MBC, the pseudorange error difference is no more than a few meters.
The overarching results show that sound speed estimates derived from eigenrays for a local grid, as opposed to a singular point, are accurate enough to support vehicle navigation.
While the NBC looks for just the least complex multipath, the high density of launch angles almost always guarantees a direct path.
Nonetheless, the consistent overestimation of pseudorange invites further analysis into acoustic arrival matching.

\subsection{Eigenray identification for beacon-to-beacon events} \label{sec:eigenrays} \llabel{1.6b}

@@

The goal of the MBC algorithm was to provide a reliable, physically intuitive interpretation of the acoustic propagation without taking on the additional burden of regularly identifying specific paths that may connect any given source-receiver pair in the network.
While it was unlikely to resolve multipath arrivals that triggered successful modem detection, an isovelocity approach would have provided no adaptivity against source and receiver depth differences.
\llabel{2.14}
Its performance was adequate for vehicle navigation and would have likely sufficed if it were not for the prominence of the duct observed relative that of other model and data products.

@@

The various sound speed inputs\textemdash isovelocity aside\textemdash not only modify the predicted effective sound speed, as seen by the colored vertical offsets, but often classify a distinct number of bounces.
HYCOM sees the most direct and one bounce multipath structures, lending a bias for faster speeds; the weighted SSP sees the most double and triple bounces, favoring slower speeds; the baseline sound speed exists in between.
Very rarely is the multipath structure classified as a direct path, where the MBC and NBC would prediction overlap.
In fact, the higher the multipath classification, the more accurate the sound speed prediction is, likely driven by a tighter or even sparser bundle of rays.
\llabel{1.19} Discontinuities in multipath classification provide initial evidence for its importance to a smoothly varying group velocity, as shown in the cluster of 30 to 30 m transmissions in \ref{fig:gvel-post}, where HYCOM jumps from one to two classified bounces amidst the baseline SSP and weighted SSP smoothly increasing while consistently seeing two and three classified bounces, respectively.
Of course, the prediction deteriorates with cross-layer transmissions across the duct, but not to the same degree at which eigenrays could not be found for the weighted SSP in section \ref{sec:eigenrays}.
The evidence suggests that the grid based method provides a useful amount of redundancy to resolve similar enough eigenrays.

It is useful to think about in what case the isovelocity\textemdash or any isovelocity framing\textemdash would have been appropriate.
The transmissions from shallow to shallow receiver would may have matched the default configuration of 1430 m/s.
The isovelocity contrived for this paper, 1441.8 m/s, best matches the transmissions from 90 to 90 m.
The one from \citet{Graupe2019}, 1450 m/s, would have had a systemic overestimation.
In addition, over the course of the four day experiment, the local maxima of the Beaufort Lens changed from roughly 1447 m/s at 40 m to 1442 m/s at 60 m.
Given that implied sound speeds just for beacon-to-beacon events span 1420 to 1445 m/s, it is safe to say that a nominal sound speed would sacrifice pseudorange accuracy somewhere, and that an adaptive approach is necessary even for short and/or small scale operations in the Beaufort Lens.

@@

\llabel{1.23} There is one example that helpfully illustrates the improvement brought upon by bounce classification.
For transmissions between North and South at 30 m, the OWTT spread is 2.1958 to 2.1963 s; the GNSS-tracked distance is 3138.54 to 3140.87 m; and the implied effective sound speed is 1429.3 to 1430.1 m/s.
For these transmissions, the weighted SSP and the MBC approach produce a pseudorange error of -1491 m, as the effective sound speed is dominated by bottom bounce arrivals with much greater travel times.
The NBC approach categorizes this same record as a quadruple surface bounce, reducing the pseudorange error to less than a meter.
Comparatively, the NBC approach for HYCOM and the baseline SSP produce pseudorange errors of 8.30 and 2.39 m, respectively. 
There is strong evidence to suggest that the sound speed and multipath fidelity codependently improve localization accuracy.

@@

\subsection{Investigating potential GNSS noise}
\llabel{1.7}
